---
title: "6_Fig6_diff_peak_annotation"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 3
    toc_float: 
      collapsed: false
      smooth_scroll: false
---

PAY attention: the big two cluster may vary (one time up down, the other time down up). so try not to run the code again as possible. Sae all the data we meed!

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
knitr::opts_knit$set(root.dir = "./CEBPA_hicut/rm_rep0/callPeak_FE4_q5_nomodel/")

```



```{r}
rm(list = ls())
PATH <- "./CEBPA_hicut/rm_rep0/callPeak_FE4_q5_nomodel/"
```




## load library
```{r}
library(ggplot2)
library(ggrepel)
library(stringr)
library(DESeq2)
library(edgeR)
library(grid)
library(factoextra)
library(cluster)
library(pbapply)
library(pheatmap)
library(openxlsx)
library(ggrepel)
library(GenomicFeatures)
library(GenomicRanges)
# library("org.Rn.eg.db") # rat
library("org.Hs.eg.db") # human
library("org.Mm.eg.db") # mm
library(TxDb.Hsapiens.UCSC.hg19.knownGene)
txdb <- TxDb.Hsapiens.UCSC.hg19.knownGene
library("biomaRt")
library(ChIPseeker)
library(ReactomePA)
library(clusterProfiler)
library(openxlsx)
library(rtracklayer)
library(plyr) # to use ioin
library(dplyr)# use full_join
library(gtools) # mixsort
library(grid)
library(futile.logger)
library(VennDiagram)

```

```{r}
rm(list = ls())
```


```{r}
plot_theme <- theme_bw() +
  theme(axis.line = element_line(colour = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(),
        # title in the middle
        plot.title = element_text(hjust = 0.5),
        text = element_text(size = 15),axis.text = element_text(size = 15)) +theme(aspect.ratio = 1)
```



```{r}
# pay attention to the / at the end 
path="./CEBPA_hicut/rm_rep0/callPeak_FE4_q5_nomodel/"

dir.create(path = paste0(path,"DiffPeak"))
dir(path) # need to produce coverage.bed and merged.bed files

```






### import peaks and coverage 

```{r}
  coverage <- read.csv(paste(path,"CEBPA_coverage.bed",sep = ''),sep='\t',header = T)
  head(coverage)
  merge <- read.csv(paste(path,"CEBPA_merged_pooled_Peaks.bed",sep = ''),sep='\t',header = F)
  names(merge)<-c('chr','start','end','idx')
  head(merge)
  tail(merge)
  
```


### set sample order
```{r}
## set sample order

extract_condition <- function(x) {return  (unlist(strsplit(x, "_rep"))[1])}
Conditions <-unique(unlist(lapply(colnames(coverage)[2:dim(coverage)[2]],extract_condition)))
Conditions


colnames(coverage)[2:dim(coverage)[2]]

sample_order <- colnames(coverage)[2:dim(coverage)[2]][c(1:3,7:9,4:6)]

length(sample_order)
sample_order

```
## re-set coverage order 
```{r}
coverage <- coverage[,c("PeakLength",sample_order)]
head(coverage)
```


## 1. PCA plot for reproducibility

####  PCA of all the samples
```{r}
RPK <- coverage[,sample_order]/coverage[,1]*10^3
TPM <- as.data.frame(t(t(RPK)/apply(RPK, 2, sum)*10^6))

pca <- as.data.frame(prcomp(t(TPM[,]),scale.=T)$x)

  pca$name <- rownames(pca);
  pca$name
  extract_condition <- function(x) {return  (unlist(strsplit(x, "[_]"))[2])}
  pca$condition  <- unlist(lapply(pca$name ,extract_condition))
  pca$condition
 
  PCA <- prcomp(t(TPM),scale.=T)
  PCA <- (PCA$sdev/sum(PCA$sdev))[1:2]
  PCA <- paste(round(PCA,4)*10^2,'%',sep = '')
  
  p<- ggplot(data = pca,aes(x=PC1, y=PC2, label=name, colour = condition)) + 
    labs(x=paste('PC1',PCA[1]),y=paste('PC2',PCA[2]))+
    geom_point() + geom_text_repel(aes(label=name), size = 4, hjust=-1, vjust=0, angle=0)+ plot_theme

  print(p)
```


```{r}
# save TPM
TPM_temp <- TPM
colnames(TPM_temp) <- paste0(colnames(TPM_temp),"_RIP_norm")

write.csv(TPM_temp,paste(path,'DiffPeak/','TPM_RIP_norm.csv',sep=''),quote = F,row.names = F,col.names = T)

```


## 2. find Diff peaks  ---heatmap on all clusters--output files

```{r}

coverage <- coverage[,-1]
head(coverage)
```

```{r}

# diff <- numeric();
Conditions 
# N = length(unique(pca$condition))
N = length(Conditions)

All_diff_Peak_idx = numeric()

for (i in 1:(N-1)){
  for (j in (i+1):N){
    # i = 1
    # j = 4
    condition1 = grep(colnames(coverage),pattern = Conditions[i])
    condition2 = grep(colnames(coverage),pattern = Conditions[j])
    
    group <- factor(c(rep(i,length(condition1)),rep(j,length(condition2))))
  
    y = DGEList(counts = coverage[,c(condition1,condition2)],group = group)
    print(names(coverage[,c(condition1,condition2)]))
    dim(y)
    
    y <- calcNormFactors(y)
    y <- estimateCommonDisp(y)
    
    fit <- glmQLFit(y)
    qlf <- glmQLFTest(fit)
    topTags(qlf)
        
    fdr <- p.adjust(p = qlf$table$PValue,method = "BH",n = length(qlf$table$PValue))
    length(fdr)
    diff_idx <- which(fdr < 0.05 & abs(qlf$table$logFC) > log2(2))
    
    ## collect all the diff peaks !
    
    All_diff_Peak_idx <- c(All_diff_Peak_idx,diff_idx)
    
    change <- cbind(merge,pval=qlf$table$PValue,qval=fdr,logFC=qlf$table$logFC)
    dim(change)
    
    print(c(paste(Conditions[i],unique(Conditions)[j],sep='<'),sum(change$logFC[diff_idx]>0),
            paste(unique(Conditions)[i],unique(Conditions)[j],sep='>'),sum(change$logFC[diff_idx]<0)))
    
    name2 <- unique(Conditions)[j]
    name1 <- unique(Conditions)[i]
    names(change)[grep(names(change),pattern="logFC")] <- paste('log2(', name2,'/', name1,')',sep = '')

    pair_name <- paste(Conditions[j],'_vs_',Conditions[i],sep='')

    dim(change)[1] == dim(coverage)[1] 
    dim(coverage)[1] ==dim(TPM)[1]


      change<- cbind(coverage[,c(condition1,condition2)],TPM[,c(condition1,condition2)],change)
    head(change)
    
      write.csv(change,paste(path,'DiffPeak/',name2,'-', name1, 'PeakChanges.csv',sep=''),quote = F,row.names = F,col.names = T)
      
      # (2)------save the diff table
      write.table(change[diff_idx,c("chr","start","end")],paste(path,'DiffPeak/', name2,'-', name1,'_diffPeak.bed',sep=''),quote = F,sep = '\t',row.names = F,col.names = F)
  
      write.csv(change[diff_idx,],paste(path,'DiffPeak/', name2,'-', name1,'_diffPeak.csv',sep=''),quote = F,row.names = F,col.names = T)
      
      ### split the file to group specific peaks
      write.table(change[diff_idx,c("chr","start","end")][which(change[diff_idx,grep(names(change),pattern="log2")] > 0 ),],paste(path,'DiffPeak/', pair_name,"_",name2,'_specific_peaks', '.bed',sep=''),quote = F,sep = '\t',row.names = F,col.names = F)
      write.table(change[diff_idx,c("chr","start","end")][which(change[diff_idx,grep(names(change),pattern="log2")]< 0 ),],paste(path,'DiffPeak/',pair_name,"_", name1,'_specific_peaks', '.bed',sep=''),quote = F,sep = '\t',row.names = F,col.names = F)
      
    ########################### ########################### ########################### ###########################

  # heat map on all the samples -------------  hcluster
  
  mtx <- TPM[diff_idx,]
  dim(mtx)
  # sample_order <- names(coverage)
    
  fun1 <- function(x) {return (substr(x,1,str_length(x)-2))}
  fun2 <- function(x) {return (substr(x,str_length(x),str_length(x)))}
  
  extract_condition <- function(x) {return  (unlist(strsplit(x, "[_]"))[2])}
  Condition <-unlist(lapply(sample_order,extract_condition))
  Condition
  Replicates <- (unlist(lapply(sample_order,fun2)))
  Replicates
  
  Sample_info <- data.frame(Conditions  = factor(Condition), Replicates  = factor(Replicates))
  ##  IMPORTANT ******* here must be consistent ********!!!!!!!
  rownames(Sample_info) <- colnames(mtx)
  Sample_info

  
  pheatmap(mtx ,cluster_rows=T,cluster_cols = F,show_rownames = F,
           scale = "row",
           # cutree_rows=5,
           angle_col='315',
           # annotation_row = gene_info , 
           annotation_col = Sample_info,
           main = paste('Diff Peak hclust: ',name2,' VS ', name1))
    
  ## other way to draw 
  # hclust -------

mtx_Zscore <- (mtx-apply(X = mtx, MARGIN = 1, mean))/apply(X = mtx, MARGIN = 1, sd)
      # remove NA or inf value  
      # rows that should kept 
      row_keep <- apply(mtx_Zscore, 1, function(x) !(sum(is.na(x)) | sum(is.infinite(x))) )
      mtx_Zscore = mtx_Zscore[row_keep, ]

mtx_hclust <- hclust(dist(mtx_Zscore), method = "complete")
      
      # peak_cluster <- data.frame(cutree(tree = as.dendrogram(mtx_hclust), k = 4)) # this is not correct,remove  as.dendrogram  (Qing)
      peak_cluster <- data.frame(cutree(tree = (mtx_hclust), k = 2))
      names(peak_cluster)<- 'cluster';peak_cluster$cluster <- factor(peak_cluster$cluster)
      pheatmap(mtx_Zscore,cluster_rows=T,cluster_cols = F,show_rownames = F,cutree_rows=2,
               angle_col='315',annotation_row = peak_cluster, annotation_col = Sample_info,main = paste('Diff Peak hclust: ',name2,' VS ', name1))
      

##  heat map on paired samples ----------kmeans

      # fviz_nbclust(mtx_Zscore, kmeans, method = "wss")
  
      set.seed(2);mtx_kmeans <- kmeans(mtx_Zscore,centers=2,nstart = 25)
      peak_kmeans <- as.data.frame(sort(mtx_kmeans$cluster)); names(peak_kmeans) <- 'cluster'; 
      peak_kmeans$cluster <- factor(peak_kmeans$cluster)
      
      pheatmap(mtx_Zscore[names(sort(mtx_kmeans$cluster)),],cluster_rows=F,cluster_cols = F,show_rownames = F,
               angle_col='315',annotation_row = peak_kmeans, annotation_col = Sample_info,main = paste('Diff Peak kmeans : ',name2,' VS ', name1))

      
  
   # heat map on paired samples -------------  hcluster
  
  mtx <- TPM[diff_idx,c(condition1,condition2)]
  dim(mtx)
  # sample_order <- names(coverage)
    
  
  
  extract_condition <- function(x) {return  (unlist(strsplit(x, "[_]"))[2])}
  Condition <-unlist(lapply(colnames(mtx),extract_condition))
  Condition
  
  fun1 <- function(x) {return  (unlist(strsplit(x, "[_]"))[3])}
  Replicates <- (unlist(lapply(colnames(mtx),fun1)))
  Replicates
  
  Sample_info <- data.frame(Conditions  = factor(Condition), Replicates  = factor(Replicates))
  ##  IMPORTANT ******* here must be consistent ********!!!!!!!
  rownames(Sample_info) <- colnames(mtx)
  Sample_info
  
  pheatmap(mtx ,cluster_rows=T,cluster_cols = F,show_rownames = F,
           scale = "row",
           # cutree_rows=5,
           angle_col='315',
           # annotation_row = gene_info , 
           annotation_col = Sample_info,
           main = paste('Diff Peak hclust: ',name2,' VS ', name1))

  ## other way to draw 
  # hclust -------

mtx_Zscore <- (mtx-apply(X = mtx, MARGIN = 1, mean))/apply(X = mtx, MARGIN = 1, sd)
      # remove NA or inf value  
      # rows that should kept 
      row_keep <- apply(mtx_Zscore, 1, function(x) !(sum(is.na(x)) | sum(is.infinite(x))) )
      mtx_Zscore = mtx_Zscore[row_keep, ]

mtx_hclust <- hclust(dist(mtx_Zscore), method = "complete")
      
      # peak_cluster <- data.frame(cutree(tree = as.dendrogram(mtx_hclust), k = 4)) # this is not correct,remove  as.dendrogram  (Qing)
      peak_cluster <- data.frame(cutree(tree = (mtx_hclust), k = 2))
      names(peak_cluster)<- 'cluster';peak_cluster$cluster <- factor(peak_cluster$cluster)
      pheatmap(mtx_Zscore,cluster_rows=T,cluster_cols = F,show_rownames = F,cutree_rows=2,
               angle_col='315',annotation_row = peak_cluster, annotation_col = Sample_info,main = paste('Diff Peak hclust: ',name2,' VS ', name1))
      

##  heat map on paired samples ----------kmeans

      # fviz_nbclust(mtx_Zscore, kmeans, method = "wss")
  
      set.seed(2);mtx_kmeans <- kmeans(mtx_Zscore,centers=2,nstart = 25)
      peak_kmeans <- as.data.frame(sort(mtx_kmeans$cluster)); names(peak_kmeans) <- 'cluster'; 
      peak_kmeans$cluster <- factor(peak_kmeans$cluster)
      
      pheatmap(mtx_Zscore[names(sort(mtx_kmeans$cluster)),],cluster_rows=F,cluster_cols = F,show_rownames = F,
               angle_col='315',annotation_row = peak_kmeans, annotation_col = Sample_info,main = paste('Diff Peak kmeans: ',name2,' VS ', name1))
  
  
  ## ------------------------ plot the result using volcano plot ------------------------
    
    head(change)
    change[["sig"]] <- as.factor(ifelse(change$idx %in% diff_idx,"Yes","No"))
    head(change)
    change[diff_idx,]
    
    p<- ggplot(data=change, aes(x=get(names(change)[grep(names(change),pattern="log2")]), y=-log10(qval))) +geom_point(col = ifelse(change$sig=="No","grey",ifelse(change[,colnames(change)[grep("log2",colnames(change))]]>0,"red","blue")), size = ifelse(change$sig=="Yes",1,0.1)) + 
      labs(y = "-log10(qval)", x = names(change)[grep(names(change),pattern="log2")],title = "Diff peaks(qval < 0.05; FC > 2)")+
       ylim(c(0, 6))+
      # add line according to the threshold use previously
        geom_hline(yintercept = -log10(0.05),col="black",lwd=0.6)+
        geom_vline(xintercept = log2(2),col="black",lwd=0.6)+
        geom_vline(xintercept = -log2(2),col="black",lwd=0.6)+plot_theme
    
    print(p)
    
    # no ylim
    p1 <- ggplot(data=change, aes(x=get(names(change)[grep(names(change),pattern="log2")]), y=-log10(qval))) + geom_point(col = ifelse(change$sig=="No","grey",ifelse(change[,colnames(change)[grep("log2",colnames(change))]]>0,"red","blue")), size = ifelse(change$sig=="Yes",1,0.1)) + 
      labs(y = "-log10(qval)", x = names(change)[grep(names(change),pattern="log2")],title = "Diff peaks(qval < 0.05; FC > 2)")+
       # add line according to the threshold use previously
        geom_hline(yintercept = -log10(0.05),col="black",lwd=0.6)+
        geom_vline(xintercept = log2(2),col="black",lwd=0.6)+
        geom_vline(xintercept = -log2(2),col="black",lwd=0.6)+plot_theme
    
    print(p1)
    
    ## compress y axis again
    p2 <-  ggplot(data=change, aes(x=get(names(change)[grep(names(change),pattern="log2")]), y=log10(-log10(qval)+1))) +geom_point(col = ifelse(change$sig=="No","grey",ifelse(change[,colnames(change)[grep("log2",colnames(change))]]>0,"red","blue")), size = ifelse(change$sig=="Yes",1,0.1)) + 
      labs(y = "log10(-log10(qval)+1)", x = names(change)[grep(names(change),pattern="log2")],title = "Diff peaks(qval < 0.05; FC > 2)")+
      # add line according to the threshold use previously
        geom_hline(yintercept = log10(-log10(0.05)+1),col="black",lwd=0.6)+
        geom_vline(xintercept = log2(2),col="black",lwd=0.6)+
        geom_vline(xintercept = -log2(2),col="black",lwd=0.6)+plot_theme
    
     print(p2)
     
     
  }
}

```



## 3. combine all the DiffPeaks 

### input all diff peaks


```{r}
path="./CEBPA_hicut/rm_rep0/callPeak_FE4_q5_nomodel/"
```

```{r}
diff_Peak_idx <- c()
files<- dir(paste0(path,"diffpeak"))[grep(dir(paste0(path,"diffpeak")),pattern = "diffPeak.csv")]
for (f in files){
  # f = files[1]
  temp <- read.csv(paste0(path,"diffpeak/",f))$idx
  
   diff_Peak_idx <- c(diff_Peak_idx,temp)
  
}
```


```{r}
length(diff_Peak_idx)
All_DiffPeaks_unique_idx <- unique(diff_Peak_idx)

length(All_DiffPeaks_unique_idx)
```
### SAVE ALL the diff peaks

```{r}
  merge <- read.csv(paste(path,"CEBPA_merged_pooled_Peaks.bed",sep = ''),sep='\t',header = F)
names(merge)<-c('chr','start','end','idx')
head(merge)

dim(merge[which(merge$idx %in% All_DiffPeaks_unique_idx),])
write.table(merge[which(merge$idx %in% All_DiffPeaks_unique_idx),],"./CEBPA_all_diffPeak.bed", row.names = F, col.names = T, quote = F, sep = "\t")
```


```{r}
TPM <- read.csv(paste(path,'DiffPeak/','TPM_RIP_norm.csv',sep=''))
dim(TPM)
```

### overall heatmap 



```{r}
# heat map on all the samples -------------hcluster
    samples <- as.data.frame(pca$condition); names(samples) <- 'sample';
    
    # rownames(samples) <- pca$name;
    rownames(samples) <- colnames(TPM) # make sure the same is consistent, because I changed the colnames of TPM previously
    samples$sample <- factor(samples$sample)
    samples

    mtx_Zscore <- (TPM[All_DiffPeaks_unique_idx,]-apply(X = TPM[All_DiffPeaks_unique_idx,], MARGIN = 1, mean))/apply(X = TPM[All_DiffPeaks_unique_idx,], MARGIN = 1, sd)
    dim(mtx_Zscore)
    # remove NA or inf value  
      # rows that should kept 
    row_keep <- apply(mtx_Zscore, 1, function(x) !(sum(is.na(x)) | sum(is.infinite(x))) )
    mtx_Zscore = mtx_Zscore[row_keep, ]
    dim(mtx_Zscore)
    
    mtx_hclust <- hclust(dist(mtx_Zscore), method = "complete")
    peak_col <- data.frame(cutree(tree = (mtx_hclust), k = 3))
    names(peak_col)<- 'cluster';peak_col$cluster <- factor(peak_col$cluster)

    pheatmap(mtx_Zscore,cluster_rows=T,cluster_cols = F,show_rownames = F,cutree_rows=2,
             angle_col='315',annotation_row = peak_col, annotation_col = samples,main = 'All DiffPeaks z-score (hclust)')
    
    
    peak_col <- data.frame(cutree(tree = (mtx_hclust), k = 4))
    names(peak_col)<- 'cluster';peak_col$cluster <- factor(peak_col$cluster)
    pheatmap(mtx_Zscore,cluster_rows=T,cluster_cols = F,show_rownames = F,cutree_rows=2,
             angle_col='315',annotation_row = peak_col, annotation_col = samples,main = 'All DiffPeaks z-score (hclust)')

```

```{r}
set.seed(2);mtx_kmeans <- kmeans(mtx_Zscore,centers=3,nstart = 25)
    peak_kmeans <- as.data.frame(sort(mtx_kmeans$cluster)); names(peak_kmeans) <- 'cluster'; 
    peak_kmeans$cluster <- factor(peak_kmeans$cluster)
    
    pheatmap(mtx_Zscore[names(sort(mtx_kmeans$cluster)),],cluster_rows=F,cluster_cols = F,show_rownames = F,
             angle_col='315',annotation_row = peak_kmeans, annotation_col = samples,main = 'All DiffPeaks z-score (k-means)')
```

```{r}
set.seed(2);mtx_kmeans <- kmeans(mtx_Zscore,centers=4,nstart = 25)
    peak_kmeans <- as.data.frame(sort(mtx_kmeans$cluster)); names(peak_kmeans) <- 'cluster'; 
    peak_kmeans$cluster <- factor(peak_kmeans$cluster)
    
    pheatmap(mtx_Zscore[names(sort(mtx_kmeans$cluster)),],cluster_rows=F,cluster_cols = F,show_rownames = F,
             angle_col='315',annotation_row = peak_kmeans, annotation_col = samples,main = 'All DiffPeaks z-score (k-means)')
```
## re-order clusters


```{r}
    # set.seed(2);mtx_kmeans <- kmeans(mtx_Zscore,centers=4,nstart = 25)
  mtx_kmeans$cluster[which(mtx_kmeans$cluster == 4)] <- "C1"
  mtx_kmeans$cluster[which(mtx_kmeans$cluster == 1)] <- "C2"
  mtx_kmeans$cluster[which(mtx_kmeans$cluster == 3)] <- "C3"
  mtx_kmeans$cluster[which(mtx_kmeans$cluster == 2)] <- "C4"


    
    gene_kmeans <- as.data.frame(sort(mtx_kmeans$cluster)); names(gene_kmeans) <- 'cluster';
    gene_kmeans$cluster <- factor(gene_kmeans$cluster)

    pheatmap(mtx_Zscore[names(sort(mtx_kmeans$cluster)),],cluster_rows=F,cluster_cols = F,show_rownames = F,
             angle_col='315',annotation_row = peak_kmeans, annotation_col = samples,main = 'All DiffPeaks z-score (k-means)')

```


```{r}
pdf("./diffpeak_4_clusters.pdf")
min(mtx_Zscore); max(mtx_Zscore)
pheatmap(mtx_Zscore[names(sort(sample(mtx_kmeans$cluster))),],cluster_rows=F,cluster_cols = F,show_rownames = F,
         breaks = seq(min(mtx_Zscore)+ 1, max(mtx_Zscore)-1,length.out = 100),
             angle_col='315',annotation_row = peak_kmeans, annotation_col = samples,main = 'All DiffPeaks z-score (k-means)')

dev.off()

```

```{r}
# names(mtx_kmeans$cluster)
    Cluster_4_list <- list()
    Cluster_4_list[["Cluster1"]]<- names(mtx_kmeans$cluster[which(mtx_kmeans$cluster=="C1")])
    Cluster_4_list[["Cluster2"]]<- names(mtx_kmeans$cluster[which(mtx_kmeans$cluster=="C2")])
    Cluster_4_list[["Cluster3"]]<- names(mtx_kmeans$cluster[which(mtx_kmeans$cluster=="C3")])
    Cluster_4_list[["Cluster4"]]<- names(mtx_kmeans$cluster[which(mtx_kmeans$cluster=="C4")])
    lapply(Cluster_4_list, length)
  
    
```






### (pay attention ) save 4 kemans clusters (also for homer analysis)
```{r}
dir.create(path = paste0(path,"DiffPeakClusters"))

write.table(merge[Cluster_4_list[["Cluster1"]],],"./CEBPA_hicut/rm_rep0/callPeak_FE4_q5_nomodel/DiffPeakClusters/CEBPA_C1_peaks.bed", row.names = F, col.names = F,quote = F, sep = "\t")

write.table(merge[Cluster_4_list[["Cluster2"]],],"./CEBPA_hicut/rm_rep0/callPeak_FE4_q5_nomodel/DiffPeakClusters/CEBPA_C2_peaks.bed", row.names = F, col.names = F,quote = F, sep = "\t")

write.table(merge[Cluster_4_list[["Cluster3"]],],"./CEBPA_hicut/rm_rep0/callPeak_FE4_q5_nomodel/DiffPeakClusters/CEBPA_C3_peaks.bed", row.names = F, col.names = F,quote = F, sep = "\t")

write.table(merge[Cluster_4_list[["Cluster4"]],],"./CEBPA_hicut/rm_rep0/callPeak_FE4_q5_nomodel/DiffPeakClusters/CEBPA_C4_peaks.bed", row.names = F, col.names = F,quote = F, sep = "\t")

```

## 4. Peak annotation

```{r}
# pay attention to the / at the end 
path="./CEBPA_hicut/rm_rep0/callPeak_FE4_q5_nomodel/"


```


## import  peaks (FE>6 q5)


```{r}

peak_list= list()


for(name in c("0h","4h","4d")){
  cat(name,"\n")
  peak_list[[name]]<- read.delim(paste0("./CEBPA_hicut/rm_rep0/callPeak_FE4_q5_nomodel/merged_macs2_no_SPMR/CEBPA_",name,"_merged_peaks.csv"),sep = ",",header = T)
  cat(dim(peak_list[[name]]),"\n")
  head(peak_list[[name]])
  # peak_list[[name]] <- peak_list[[name]][which(peak_list[[name]]$fold_enrichment >= 6 & peak_list[[name]]$X.NAME..1 >=5),]
  # 
  cat(dim(peak_list[[name]]),"\n")
  cat("\n")
  peak_list[[name]]<- GRanges(peak_list[[name]])
}


```





```{r}

lapply(peak_list, length)

```

## peak annotation
```{r}
txdb <- TxDb.Hsapiens.UCSC.hg19.knownGene
peakAnnoList <- lapply(peak_list, annotatePeak, TxDb=txdb, tssRegion=c(-1000, 1000), verbose=FALSE, annoDb="org.Hs.eg.db")
```

## save 
```{r}

for (C in names(peakAnnoList)){
  # C ="0h" 
  temp <- as.data.frame(peakAnnoList[[C]]@anno)

  write.csv(temp,paste0("./merged_macs2_no_SPMR/CEBPA_",C,"_Peak_FE6q5_aanotation.csv"))
}




```





```{r}
plotAnnoBar(peakAnnoList)

names(peakAnnoList)

for (c in names(peakAnnoList)){
  plotAnnoPie(peakAnnoList[[c]])
  peakAnnoList[[c]]
}

```

### Profile  to TSS regions
```{r}

promoter <- getPromoters(TxDb=txdb, upstream=1000, downstream=1000)

tagMatrixList <- lapply(peak_list, getTagMatrix, windows=promoter) 


## Average Profile -----
plotAvgProf(tagMatrixList, xlim=c(-1000, 1000), xlab="Genomic Region (5'->3')",
        ylab = "Read Count Frequency", facet="row")
# put together, no facet 
plotAvgProf(tagMatrixList, xlim=c(-1000, 1000), xlab="Genomic Region (5'->3')",
        ylab = "Read Count Frequency")

```





## 5. GO analysis

ClusterProfiler(ORA)---DiffPeak---GO and KEGG analysis


#### fun_ClusterProfiler_GO_KEGG_human
```{r}
## --2.-------Run GO and KEGG analysis -----------
fun_ClusterProfiler_GO_KEGG_human <- function(gene_ENTREZID,Gene){

  # gene_ENTREZID <- IDs
  #      1. GO analysis ----- enrichGO 
  ego_all <- enrichGO(gene = gene_ENTREZID,
                      # pay attention here----- use the correct ref
                      # OrgDb="org.Mm.eg.db",
                      OrgDb="org.Hs.eg.db",
                      ont='ALL',pAdjustMethod = 'BH',
                      pvalueCutoff = 1,  qvalueCutoff = 1,
                      # pvalueCutoff = 0.05, qvalueCutoff = 0.2,
                      keyType = 'ENTREZID')
  
  pdf_name <- paste(Gene,"_GO_KEGG_figures.pdf",sep ="_")
  pdf(pdf_name)
  p_dot_GO <- dotplot(ego_all,showCategory=20,title =paste(Gene,"(GO)",sep =""))
  print(p_dot_GO)
  ego_all_sig <- ego_all[ego_all$pvalue <= 0.05]
  write.table(ego_all,file = paste(Gene,"GOenrich.txt",sep ="_"),sep = "\t")
  write.table(ego_all_sig,file = paste(Gene,"GOenrich_significant.txt",sep ="_"),sep = "\t")
  
  p_all<- read.table( paste(Gene,"GOenrich.txt",sep ="_"))
  p_all<-p_all[order(p_all$p.adjust),]
  dim(p_all)
  head(p_all)
  p_BP<-p_all[which(p_all[,1]== "BP"),]
  write.table(p_BP,file = paste(Gene,"BP_GOenrich.txt",sep ="_"),sep = "\t")
  p_CC<-p_all[which(p_all[,1]== "CC"),]
  write.table(p_CC,file = paste(Gene,"CC_GOenrich.txt",sep ="_"),sep = "\t")
  p_MF<-p_all[which(p_all[,1]== "MF"),]
  write.table(p_MF,file = paste(Gene,"MF_GOenrich.txt",sep ="_"),sep = "\t")
  ####    customize the bar plot 
  ## all term
  par(mar=c(4,20,2,2)+0.1)
  # down left up right
  barplot(rev(-log10(p_all$p.adjust[1:20]) ),horiz=TRUE,main = paste(Gene,"(GO)",sep = ""),col = "#0080CC",
          names.arg = rev(paste(p_all$Description,"(",p_all$Count,")",sep ="")[1:20]),las =1,
          xlim = c(0,max(rev(-log10(p_all$p.adjust[1:20]))+0.5)),
          xlab = "-log10(p.adjust)" )
  # paste(p_all$Description,"(",p_all$Count,")",sep ="")
  
  bar_plot_space <- seq(from =0,by = 0.2,length.out=20)
  x_pos <- rev(-log10(p_all$p.adjust[1:20]))+0.2
  y_pos = 1:20 + bar_plot_space
  ## padj = 0.05
  index <-which(x_pos>1.3)
  points(x= x_pos[index],y=y_pos[index],pch  = "*")
  
  ## BP CC MF term
  
  par(mar=c(4,20,2,2)+0.1)
  barplot(rev(-log10(p_BP$p.adjust[1:20]) ),horiz=TRUE,main = paste(Gene,"(BP)",sep = ""),col = "#0080CC",
          names.arg = rev(paste(p_BP$Description,"(",p_BP$Count,")",sep ="")[1:20]),las =1,
          xlim = c(0,max(rev(-log10(p_BP$p.adjust[1:20]))+0.5)),
          xlab = "-log10(p.adjust)")
  bar_plot_space <- seq(from =0,by = 0.2,length.out=20)
  x_pos <- rev(-log10(p_BP$p.adjust[1:20]))+0.2
  y_pos = 1:20 + bar_plot_space
  ## padj = 0.05
  index <-which(x_pos>1.3)
  points(x= x_pos[index],y=y_pos[index],pch  = "*")
  
  par(mar=c(4,20,2,2)+0.1)
  # down left up right
  barplot(rev(-log10(p_CC$p.adjust[1:20]) ),horiz=TRUE,main = paste(Gene,"(CC)",sep = ""),col = "#0080CC",
          names.arg = rev(paste(p_CC$Description,"(",p_CC$Count,")",sep ="")[1:20]),las =1,
          xlim = c(0,max(rev(-log10(p_CC$p.adjust[1:20]))+0.5)),
          xlab = "-log10(p.adjust)")
  bar_plot_space <- seq(from =0,by = 0.2,length.out=20)
  x_pos <- rev(-log10(p_CC$p.adjust[1:20]))+0.2
  y_pos = 1:20 + bar_plot_space
  ## padj = 0.05
  index <-which(x_pos>1.3)
  points(x= x_pos[index],y=y_pos[index],pch  = "*")
  
  par(mar=c(4,20,2,2)+0.1)
  # down left up right
  barplot(rev(-log10(p_MF$p.adjust[1:20]) ),horiz=TRUE,main = paste(Gene,"(MF)",sep = ""),col = "#0080CC",
          names.arg = rev(paste(p_MF$Description,"(",p_MF$Count,")",sep ="")[1:20]),las =1,
          xlim = c(0,max(rev(-log10(p_MF$p.adjust[1:20]))+0.5)),
          xlab = "-log10(p.adjust)")
  bar_plot_space <- seq(from =0,by = 0.2,length.out=20)
  x_pos <- rev(-log10(p_MF$p.adjust[1:20]))+0.2
  y_pos = 1:20 + bar_plot_space
  ## padj = 0.05
  index <-which(x_pos>1.3)
  points(x= x_pos[index],y=y_pos[index],pch  = "*")
  
  # 2. KEGG analysis  
  
  ekegg <- enrichKEGG(gene_ENTREZID, organism = 'hsa', keyType = 'kegg',
                      # -------------human : hsa  Rat:rno    Mouse:mmu-------------------
                      # pvalueCutoff = 0.05,qvalueCutoff = 0.2,
                      pvalueCutoff = 1,  qvalueCutoff = 1,
                      pAdjustMethod = 'BH', minGSSize = 10,maxGSSize = 500,
                      use_internal_data = FALSE)
  p_dot_KEGG<- dotplot(ekegg, showCategory=30,title = paste(Gene,"(KEGG)",sep = "") )
  print(p_dot_KEGG)
  
  dev.off()
}

```

#### Peaks associated genes
```{r}
diff_peak_gene_list <- list()

for(name in names(peakAnnoList)){
  print(name)
  diff_peak_gene_list[[name]] <- unique(as.data.frame(peakAnnoList[[name]])["geneId"])

}

lapply(diff_peak_gene_list, dim)

```

####  analysis 

```{r}
for(name in names(diff_peak_gene_list)){
  
  print(name)
  fun_ClusterProfiler_GO_KEGG_human(diff_peak_gene_list[[name]]$geneId,name)
}



```


